---
title: "607_assignment_5"
author: "Eddie Xu"
date: "2023-02-24"
output: html_document
---

```{r}
library(tidyverse)
```

## Data 607 Assignment 5
Given a messy data, the goal is to tidy and transform data to suit the need for data analysis.

## Read CSV
```{r}
readfile <- data.frame(read.csv("https://raw.githubusercontent.com/eddiexunyc/607_assignment_5/main/tidy%20data.csv", header =TRUE))

head(readfile)
```

## Data Cleaning
Since the csv file is untidy and does not follow the principle of tidy data where:

1) Every column is a variable
2) Every row is an observation
3) Every cell is a single value

The data will be cleansed to reflect that.
```{r}
##split the data based on the row number
alaska_df <- readfile[1:2,]
amwest_df <- readfile[4:5,]
```
```{r}
##rename the column 
colnames(alaska_df) <- c("Airline", "Arrival Status", "Los Angeles", "Phoenix", "San Diego","San Francisco", "Seattle")

##fill in the null value with new value for Alaska
alaska_df$Airline[alaska_df$Airline == ""] <-c("ALASKA")

##pivot longer the city and value
alaska_tidy_data <- alaska_df %>%
  pivot_longer(cols = c("Los Angeles", "Phoenix", "San Diego","San Francisco", "Seattle"), names_to = "City", values_to = "Count")

knitr::kable(alaska_tidy_data)
```


```{r}
##rename the column
colnames(amwest_df) <- c("Airline", "Arrival Status", "Los Angeles", "Phoenix", "San Diego","San Francisco", "Seattle")

##fill in the null value with new value
amwest_df$Airline[amwest_df$Airline == ""] <-c("AM WEST")

##pivot longer the city and value
amwest_tidy_data <- amwest_df %>%
  pivot_longer(cols = c("Los Angeles", "Phoenix", "San Diego","San Francisco", "Seattle"), names_to = "City", values_to = "Count")

knitr::kable(amwest_tidy_data)
```

```{r}
##combine both amwest and alaska dataframe together
airline_tidy_data <- rbind(alaska_tidy_data, amwest_tidy_data)

knitr::kable(airline_tidy_data)
```

##Data Analysis
Now that the data is in a tidy format, the data analysis can be done on the airline data.

```{r}
##summarize the data
airline_summary <- airline_tidy_data %>% group_by(`Airline`, `Arrival Status`) %>% 
  summarize_at('Count', mean)

head(airline_summary)
```

Let's see the arrival status in a bar graph
```{r}
##bar graph based on the arrival status by city.

ggplot(airline_tidy_data, aes(City, y= Count, fill= `Arrival Status`)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r}
```